% ====================================================================
%  Acknowledgements — place at the end of the document
% ====================================================================

\newpage
\section*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}
\hskip \parindent
\par These notes owe their existence to three principal sources: the lectures
delivered by \textbf{Prof.\ Andrea Solotar} in the course 104168 Algebra~B
at Guangdong Technion --- Israel Institute of Technology, the accompanying
tutorial materials prepared by \textbf{Carlos Antunes Percincula}, and the
textbook \textit{Linear Algebra} (2nd~ed.)\ by K.~Hoffman and R.~Kunze.
The clarity and depth of these sources made the present compilation
possible.

\par It should be noted, however, that the presentation herein departs
significantly from the original materials.  The content has been
restructured to improve logical coherence, augmented with supplementary
definitions, examples, and complete proofs where the originals were only
sketched, and in places rewritten for precision and readability.  Any
errors or inaccuracies introduced during this process remain the sole
responsibility of the author.

\par Special thanks are due to \textbf{Congyi Mai}, whose thorough proofreading
and insightful comments led to numerous improvements throughout the
manuscript.

\vspace{1.5em}
\begin{flushright}
\textit{Xiangjun (Samuel) Liu}\\[2pt]
Spring Semester, 2025--2026
\end{flushright}

\vspace{1em}
\noindent\rule{\textwidth}{0.4pt}
\smallskip
\noindent{\small\textcopyright\ 2025--2026 Xiangjun Liu. All rights reserved.}
