% ====================================================================
%  Acknowledgements
% ====================================================================

\newpage
\section*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}
\hskip \parindent
\par These notes are a restructured and reformatted edition of the original
\textit{Notes of 104016 Algebra~1/Extended} by the same author, adapted
to the \texttt{tcolorbox} visual framework for improved readability and
navigability.  The mathematical content is based on three principal
sources: the blackboard lectures of \textbf{Cristian Chaparro} in the
course 104016 Algebra~1/Extended at Guangdong Technion --- Israel
Institute of Technology, the accompanying problem sessions and workshop
materials, and the textbook \textit{Linear Algebra} (2nd~ed.)\ by
K.~Hoffman and R.~Kunze.  The clarity and rigour of these sources made
the present compilation possible.

\par It should be noted that the presentation herein departs from the
original materials in several respects.  The content has been restructured
to improve logical coherence and augmented with supplementary definitions,
examples, complete proofs, warning boxes highlighting notation differences,
and \texttt{tcolorbox} environments that visually distinguish definitions,
theorems, propositions, corollaries, proofs, examples, remarks, and
warnings.  Any errors or inaccuracies introduced during this process remain
the sole responsibility of the author.

\par Parts of this document, including but not limited to sections on finite
fields and modular arithmetic, are adapted from or based on content from
\textbf{Wikipedia}, which is licensed under the Creative Commons
Attribution--ShareAlike 4.0 International License.  The original source is
available at \url{https://www.wikipedia.org/}.  The adapted material has
been reorganised, rewritten, and integrated into these notes.  In
accordance with the ShareAlike requirement, this document is distributed
under the \textbf{CC BY-SA 4.0} license.

\vspace{1.5em}
\begin{flushright}
\textit{Xiangjun (Samuel) Liu}\\[2pt]
2025--2026
\end{flushright}

\vspace{1em}
\noindent\rule{\textwidth}{0.4pt}
\smallskip
\noindent{\small\textcopyright\ 2025--2026 Xiangjun Liu.
Licensed under CC BY-SA 4.0.}
